---
title: "List of Unicorn Companies Analysis"
author: "Wlademir Ribeiro Prates"
date: "2022-07-29"
categories: [Reporting]
---

The full and updated list of Unicorn companies used on this analysis can be found [here](https://www.cbinsights.com/research-unicorn-companies), at the CBInsights website.

## Reading Unicorn companies list

First, we need to load the used packages and read the data. Below I am showing the top 5 unicorn companies, with all the variables from the raw data.

```{r}
#| warning: false
#| message: false

# Loading packages
library(dplyr)
library(echarts4r)
library(glue)
library(lubridate)
library(reactable)
library(readxl)

# Reading the Excel file
unicorn_data <- readxl::read_excel("unicorn_companies.xlsx", skip = 2) |>
  dplyr::arrange(dplyr::desc("Valuation ($B)")) |>
  dplyr::mutate(
    `Years as Unicorn` = round(
      (lubridate::interval(`Date Joined`, lubridate::today()) %/% months(1)) / 12,
      digits = 2
    ),
    `Number of Investors` = stringr::str_count(`Select Investors`, ',') + 1,
    `Industry` = as.factor(`Industry`)
  )

interactive_table_data <- unicorn_data |>
  dplyr::select(-`Date Joined`) |>
  dplyr::mutate(`Years as Unicorn` = floor(`Years as Unicorn`))

# Creating an interactive table
reactable::reactable(
  interactive_table_data,
  columns = list(
    `Select Investors` = colDef(minWidth = 180),
    Country = colDef(html = TRUE)
  ),
  style = list(fontSize = "0.70rem"),
  defaultPageSize = 10,
  searchable = TRUE,
  striped = TRUE,
  resizable = TRUE
)
```

## Treemap

```{r}
tree_data <- unicorn_data |>
  dplyr::arrange(dplyr::desc(`Valuation ($B)`)) |>
  dplyr::select(Industry, name = Company, value = `Valuation ($B)`) |>
  tidyr::nest(children = c("name", "value")) |>
  dplyr::mutate(value = purrr::map_dbl(children, ~ sum(.x$value))) |>
  dplyr::mutate(children = purrr::map(children, ~ filter(., value >= 5))) |>
  dplyr::rename(name = Industry)

tree_data |>
  echarts4r::e_charts() |>
  echarts4r::e_title(text = "Unicorn companies by Industry") |>
  echarts4r::e_theme("infographic") |>
  echarts4r::e_treemap(
    leafDepth = 2,
    breadcrumb = TRUE,
    backgroundColor = "red",
    itemStyle = list(normal = list(
      borderWidth = 0,
      gapWidth = 2,
      backgroundColor = "gray70"
    )),
    name = "Having fun with treemaps",
    upperLabel = list(normal = list(
      show = TRUE,
      height = 30,
      formatter = "{b}",
      color = "black",
      fontSize = 12
    ))
  ) |>
  echarts4r::e_tooltip()
```


## Boxplot by Industry

```{r}
boxplot_by_category <- function(data, category, outliers) {
  data |>
    dplyr::group_by(!!rlang::sym(category)) |>
    dplyr::mutate(median = median(`Valuation ($B)`)) |>
    dplyr::arrange(dplyr::desc(median)) |>
    echarts4r::e_charts() |>
    echarts4r::e_theme("essos")  |>
    echarts4r::e_color(background = "white") |>
    echarts4r::e_boxplot(serie = `Valuation ($B)`, outliers = outliers) |>
    echarts4r::e_tooltip(trigger = "item") |>
    echarts4r::e_datazoom() |>
    echarts4r::e_title(glue::glue("Valuation ($B) Boxplot by {category}")) |>
    echarts4r::e_y_axis(type = "log")
}

boxplot_by_category(unicorn_data, category = "Industry", outliers = TRUE)
```

### Scatterplot

```{r}
unicorn_scatterplot <- function(data) {
  data |>
    echarts4r::e_charts(`Years as Unicorn`) |>
    echarts4r::e_scatter(`Valuation ($B)`, `Number of Investors`, bind = Company) |>
    echarts4r::e_labels(show = FALSE, position = "top") |>
    echarts4r::e_title(glue::glue("Valuation ($B) vs Years as Unicorn vs Number of Investors Scatterplot")) |>
    echarts4r::e_y_axis_(type = "log") |>
    echarts4r::e_tooltip(
      axisPointer = list(type = "cross"),
      formatter = htmlwidgets::JS("
      function(params){return(
        '<strong>' + params.name + '</strong>' + '<br>' +
          'Industry: <i>' + params.seriesName + '</i><br>' +
          'Valuation ($B): <i>' + parseFloat((params.value[1] * 10) / 10).toFixed(1) + '</i><br>' +
          'Years as Unicorn: <i>' + params.value[0] + '</i><br>' +
          'Number of Investors: <i>' + params.value[2] + '</i>'
      )}"
      )
    ) |>
    echarts4r::e_axis_labels(x = "Years as Unicorn", y = "Valuation ($B) - log scale") |>
    echarts4r::e_legend(show = TRUE, type = "scroll", right = "0", left = "0", bottom = 1) |>
    echarts4r::e_dims(height = "auto", width = "auto")
}
```

::: {.panel-tabset}

### Tech

```{r}
tech <- c(
  "Artificial intelligence",
  "Fintech",
  "Data management & analytics",
  "Edtech",
  "Other"
)
unicorn_data |>
  dplyr::filter(Industry %in% tech) |>
  dplyr::group_by(Industry) |>
  unicorn_scatterplot()
```


### Sales

```{r}
sales <- c(
  "E-commerce & direct-to-consumer",
  "Consumer & retail"
)
unicorn_data |>
  dplyr::filter(Industry %in% sales) |>
  dplyr::group_by(Industry) |>
  unicorn_scatterplot()
```


### Movement

```{r}
movement <- c(
  "Supply chain, logistics, & delivery",
  "Auto & transportation",
  "Travel",
  "Health"
)
unicorn_data |>
  dplyr::filter(Industry %in% movement) |>
  dplyr::group_by(Industry) |>
  unicorn_scatterplot()
```


### Informatics

```{r}
informatics <- c(
  "Internet software & services",
  "Cybersecurity",
  "Hardware",
  "Mobile & telecommunications"
)
unicorn_data |>
  dplyr::filter(Industry %in% informatics) |>
  dplyr::group_by(Industry) |>
  unicorn_scatterplot()
```

:::

Para investigar:

- X empresas equivalem a X% de todo o valuation das Unicorns: Gráfico de linha com soma cumulativo do valuation 
- Boxplot por grupamentos: country, industry, intervalo de tempo de vida da empresa
